---
title: "Biochar Consulting"
author: "Emily Mai"
date: "4/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# download necessary libraries
# install.packages("rstatix")
# install.packages("ggplot2")
# install.packages("gridExtra")
# install.packages("dae")
# install.packages("ggpubr")
```

```{r, include=FALSE}
library(ggplot2)
library(dplyr)
# library(rstatix)
library(readxl)
# library(reshape2)
library(gridExtra)
# library(ggpubr)
library(MASS)
library(dae)
library(emmeans)
```

## Read and format data
```{r}
# read in data
bc_data <- read_excel("C:/Users/Emily Mai/Downloads/biochar study data.xls")

# rename columns for ease of access
bc_data <- bc_data %>%
  rename(shoot_weight = `shoot weight (g)`,
         shallow_root_weight = `shallow root weight (g)`,
         deep_root_weight = `deep root weight (g)`,
         total_root_weight = `total root weight`,
         root_shoot_ratio = `root:shoot ratio`) %>%
  na.omit()

attach(bc_data)
names(bc_data)
bc_data <- subset(bc_data, select = -c(rep, type))
```

## Box Plots for Shoot Weight
```{r}
plot_shoot_weight <- bc_data[c("Wheat","Soil","Biochar","shoot_weight")]

g1 <- ggplot(data=plot_shoot_weight) +
  geom_boxplot(aes(x=Wheat, y=shoot_weight, fill=Wheat)) +
  scale_fill_manual(values=c("tan1", "steelblue2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g2 <- ggplot(data=plot_shoot_weight) +
  geom_boxplot(aes(x=Soil, y=shoot_weight, fill=Soil)) +
  scale_fill_manual(values=c("cornsilk", "plum2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g3 <- ggplot(data=plot_shoot_weight) +
  geom_boxplot(aes(x=Biochar, y=shoot_weight, fill=Biochar)) +
  scale_fill_manual(values=c("bisque", "cyan3")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())

grid.arrange(g1, g2, g3, nrow=1, top="Box Plot of Shoot Weight")
```

# Shoot Weight
NOTE: All models and analysis performed below is for shoot weight only

## Fit Full Model
```{r}
# fit full model
SW_model <- aov(shoot_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SW_model)
```

### Perform Box-Cox Transformation
```{r}
T <- boxcox(shoot_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
lambda <- T$x[which(T$y == max(T$y))]
```

### Refit and Check Assumptions after Transformation
```{r, message=F}
shoot_weight_t <- shoot_weight**lambda

# refit model
SW_model_new <- aov(shoot_weight_t ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)

# check assumptions
plot(SW_model_new)
```

### Check for 3-way Interactions
```{r}
anova(SW_model_new)
```
**Remove 3 way interaction**

## Fit Model 2
```{r}
# fit full model
SW_model2 <- aov(shoot_weight_t ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SW_model2)
```

### Check for Nonsignificant Interactions
```{r}
anova(SW_model2)
```
**Remove Wheat:Biochar interaction**

## Fit Model 3
```{r}
# fit full model
SW_model3 <- aov(shoot_weight_t ~ Wheat + Soil + Biochar + Wheat*Soil + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SW_model3)
```

### Check for Nonsignificant Interactions
```{r}
anova(SW_model3)
```
**Remove Soil:Biochar interaction**

## Fit Model 4
```{r}
# fit full model
SW_model4 <- aov(shoot_weight_t ~ Wheat + Soil + Biochar + Wheat*Soil, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SW_model4)
```

### Check for Nonsignificant Interactions
```{r}
anova(SW_model4)
```
**Remove Biochar**

## Fit Model 5
```{r}
# fit full model
SW_model5 <- aov(shoot_weight_t ~ Wheat + Soil + Wheat*Soil, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SW_model5)
```

### Check for Nonsignificant Interactions
```{r}
anova(SW_model5)
```
**Final Model: Order 2 with interaction**

## Final Model: Shoot Weight = Wheat + Soil + Wheat*Soil
```{r}
interaction.plot(Soil, Wheat, shoot_weight_t)
```

### Multiple Comparison
```{r}
# compare the effect of Wheat at each level of Soil 
emmeans(SW_model5, list(pairwise ~ Wheat|Soil)) 
```
+ The effect of wheat is significant when soil type is "soil-sand" (p-value = 0.0012 < 0.05) in terms of shoot weight.

```{r}
# compare the effect of Soil at each level of Wheat
emmeans(SW_model5, list(pairwise ~ Soil|Wheat))
```
+ The effect of Soil is significant at each level of wheat type (both p-value < 0.0001) in terms of shoot weight.

# Shallow Root Weight
NOTE: All models and analysis performed below is for shoot weight only

## Box Plot Shallow Root Weight
```{r}
plot_shallow_root_weight <- bc_data[c("Wheat","Soil","Biochar","shallow_root_weight")]
gg_object <- ggplot(data=plot_shallow_root_weight)

g1 <- gg_object +
  geom_boxplot(aes(x=Wheat, y=shallow_root_weight, fill=Wheat)) +
  scale_fill_manual(values=c("tan1", "steelblue2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g2 <- gg_object +
  geom_boxplot(aes(x=Soil, y=shallow_root_weight, fill=Soil)) +
  scale_fill_manual(values=c("cornsilk", "plum2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g3 <- gg_object +
  geom_boxplot(aes(x=Biochar, y=shallow_root_weight, fill=Biochar)) +
  scale_fill_manual(values=c("bisque", "cyan3")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())

grid.arrange(g1, g2, g3, nrow=1, top="Box Plot of Shallow Root Weight")
```

## Fit Full Model
```{r}
# fit full model
SRW_model <- aov(shallow_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SRW_model)
```
**Transformation not necessary**

### Check for Nonsignificant Interactions
```{r}
anova(SRW_model)
```
**Remove 3 way interaction**

## Fit Model 2
```{r}
# fit full model
SRW_model2 <- aov(shallow_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SRW_model2)
```

### Check for Nonsignificant Interactions
```{r}
anova(SRW_model2)
```
**Remove Wheat:Biochar interaction**

## Fit Model 3
```{r}
# fit full model
SRW_model3 <- aov(shallow_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SRW_model3)
```

### Check for Nonsignificant Interactions
```{r}
anova(SRW_model3)
```
**Remove Wheat:Soil interaction**

## Fit Model 4
```{r}
# fit full model
SRW_model4 <- aov(shallow_root_weight ~ Wheat + Soil + Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SRW_model4)
```

### Check for Nonsignificant Interactions
```{r}
anova(SRW_model4)
```
**Remove Soil:Biochar**

## Fit Model 5
```{r}
# fit full model
SRW_model5 <- aov(shallow_root_weight ~ Wheat + Soil + Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SRW_model5)
```

### Check for Nonsignificant Interactions
```{r}
anova(SRW_model5)
```
**remove Wheat**

## Fit Model 6
```{r}
# fit full model
SRW_model6 <- aov(shallow_root_weight ~ Soil + Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(SRW_model6)
```

### Check for Nonsignificant Interactions
```{r}
anova(SRW_model6)
```
**Final Model: Order 2 Model**

## Final Model: Shallow Root Weight = Soil + Biochar

### Compare main Effects
```{r}
TukeyHSD(SRW_model6)
```
```{r}
emmeans(SRW_model6,list(pairwise ~ Soil))
```
```{r}
emmeans(SRW_model6,list(pairwise ~ Biochar))
```

# Deep Root Weight
NOTE: All models and analysis performed below is for deep root weight only

## Box Plot for Deep Root Weight
```{r}
plot_deep_root_weight <- bc_data[c("Wheat","Soil","Biochar","deep_root_weight")]
gg_object <- ggplot(data=plot_deep_root_weight)

g1 <- gg_object +
  geom_boxplot(aes(x=Wheat, y=deep_root_weight, fill=Wheat)) +
  scale_fill_manual(values=c("tan1", "steelblue2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g2 <- gg_object +
  geom_boxplot(aes(x=Soil, y=deep_root_weight, fill=Soil)) +
  scale_fill_manual(values=c("cornsilk", "plum2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g3 <- gg_object +
  geom_boxplot(aes(x=Biochar, y=deep_root_weight, fill=Biochar)) +
  scale_fill_manual(values=c("bisque", "cyan3")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())

grid.arrange(g1, g2, g3, nrow=1, top="Box Plot of Deep Root Weight")
```

## Fit Full Model
```{r}
# fit full model
DRW_model <- aov(deep_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model)
```
**Transformation is necessary**

### Perform Box-Cox Transformation
```{r}
T1 <- boxcox(deep_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
lambda1 <- T1$x[which(T1$y == max(T1$y))]
```

### Refit and Check Assumptions after Transformation
```{r, message=F}
deep_root_weight_t <- deep_root_weight**lambda1

# refit model
DRW_model_new <- aov(deep_root_weight_t ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)

# check assumptions
plot(DRW_model_new)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model_new)
```
**Remove 3 way interaction**

## Fit Model 2
```{r}
# fit full model
DRW_model2 <- aov(deep_root_weight_t~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model2)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model2)
```
**Remove Wheat:Biochar interaction**

## Fit Model 3
```{r}
# fit full model
DRW_model3 <- aov(deep_root_weight_t ~ Wheat + Soil + Biochar + Wheat*Soil + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model3)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model3)
```
**Remove Wheat:Soil interaction**

## Fit Model 4
```{r}
# fit full model
DRW_model4 <- aov(deep_root_weight_t ~ Wheat + Soil + Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model4)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model4)
```
**Remove Wheat**

## Fit Model 5
```{r}
# fit full model
DRW_model5 <- aov(deep_root_weight_t ~ Soil + Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model5)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model5)
```
**remove soil:biochar interaction**

## Fit Model 6
```{r}
# fit full model
DRW_model6 <- aov(deep_root_weight_t ~ Soil + Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model6)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model6)
```
**remove soil**

## Fit Model 7
```{r}
# fit full model
DRW_model7 <- aov(deep_root_weight_t ~ Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(DRW_model7)
```

### Check for Nonsignificant Interactions
```{r}
anova(DRW_model7)
```
**Final Model: One-Way**

## Final Model: Deep Root Weight = Biochar

### Compare main effect
```{r}
TukeyHSD(DRW_model7)
```
```{r}
emmeans(DRW_model7,list(pairwise ~ Biochar))
```


# Total Root Weight
NOTE: All models and analysis performed below is for total root weight only

## Box Plot for Total Root Weight
```{r}
plot_total_root_weight <- bc_data[c("Wheat","Soil","Biochar","total_root_weight")]
gg_object <- ggplot(data=plot_total_root_weight)

g1 <- gg_object +
  geom_boxplot(aes(x=Wheat, y=total_root_weight, fill=Wheat)) +
  scale_fill_manual(values=c("tan1", "steelblue2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g2 <- gg_object +
  geom_boxplot(aes(x=Soil, y=total_root_weight, fill=Soil)) +
  scale_fill_manual(values=c("cornsilk", "plum2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g3 <- gg_object +
  geom_boxplot(aes(x=Biochar, y=total_root_weight, fill=Biochar)) +
  scale_fill_manual(values=c("bisque", "cyan3")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())

grid.arrange(g1, g2, g3, nrow=1, top="Box Plot of Total Root Weight")
```

## Fit Full Model
```{r}
# fit full model
TRW_model <- aov(total_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(TRW_model)
```

### Check for 3-way Interactions
```{r}
anova(TRW_model)
```
**Remove 3 way interaction**

## Fit Model 2
```{r}
# fit full model
TRW_model_2 <- aov(total_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(TRW_model_2)
```

### Check for Nonsignificant Interactions
```{r}
anova(TRW_model_2)
```
**Remove Wheat:Biochar interaction**

## Fit Model 3
```{r}
# fit full model
TRW_model_3 <- aov(total_root_weight ~ Wheat + Soil + Biochar + Wheat*Soil + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(TRW_model_3)
```

### Check for Nonsignificant Interactions
```{r}
anova(TRW_model_3)
```
**Remove Wheat:Soil interaction**

## Fit Model 4
```{r}
# fit full model
TRW_model_4 <- aov(total_root_weight ~ Wheat + Soil + Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(TRW_model_4)
```

### Check for Nonsignificant Interactions
```{r}
anova(TRW_model_4)
```
**Remove Wheat**

## Fit Model 5
```{r}
# fit full model
TRW_model_5 <- aov(total_root_weight ~ Soil + Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(TRW_model_5)
```

### Check for Nonsignificant Interactions
```{r}
anova(TRW_model_5)
```
**Final Model: Order 2 Model with interaction **

## Final Model: Total Root Weight = Soil + Biochar + Soil:Biochar
```{r}
interaction.plot(Biochar, Soil, total_root_weight)
```

### Multiple Comparison
```{r}
# compare the effect of Wheat at each level of Soil 
emmeans(TRW_model_5, list(pairwise ~ Soil|Biochar)) 
```
+ The effect of Soil is significant when Biochar treatment is "yes" (p-value = 0.0088 < 0.05) in terms of total root weight.

```{r}
# compare the effect of Soil at each level of Wheat
emmeans(TRW_model_5, list(pairwise ~ Biochar|Soil))
```
+ The effect of Biochar is significant when Soil type is "soil" (p-value = 0.0009 < 0.05) in terms of total root weight.

# Root Shoot Ratio
NOTE: All models and analysis performed below is for root shoot ratio only

## Box Plot for Root Shoot Ratio
```{r}
plot_root_shoot_ratio <- bc_data[c("Wheat","Soil","Biochar","deep_root_weight")]
gg_object <- ggplot(data=plot_root_shoot_ratio)

g1 <- gg_object +
  geom_boxplot(aes(x=Wheat, y=root_shoot_ratio, fill=Wheat)) +
  scale_fill_manual(values=c("tan1", "steelblue2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g2 <- gg_object +
  geom_boxplot(aes(x=Soil, y=root_shoot_ratio, fill=Soil)) +
  scale_fill_manual(values=c("cornsilk", "plum2")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())
g3 <- gg_object +
  geom_boxplot(aes(x=Biochar, y=root_shoot_ratio, fill=Biochar)) +
  scale_fill_manual(values=c("bisque", "cyan3")) +
  theme(axis.text.x = element_blank(),axis.ticks = element_blank())

grid.arrange(g1, g2, g3, nrow=1, top="Box Plot Root Shoot Ratio")
```

## Fit Full Model
```{r}
# fit full model
RSR_model <- aov(root_shoot_ratio ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar + Wheat*Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(RSR_model)
```

### Check for 3-way Interactions
```{r}
anova(RSR_model)
```
**Remove 3 way interaction**

## Fit Model 2
```{r}
# fit full model
RSR_model_2 <- aov(root_shoot_ratio ~ Wheat + Soil + Biochar + Wheat*Soil + Wheat*Biochar + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(RSR_model_2)
```

### Check for Nonsignificant Interactions
```{r}
anova(RSR_model_2)
```
**Remove Wheat:Biochar interaction**

## Fit Model 3
```{r}
# fit full model
RSR_model_3 <- aov(root_shoot_ratio ~ Wheat + Soil + Biochar + Wheat*Soil + Soil*Biochar, data=bc_data)
```

### Check Model Assumptions
```{r, message=F}
# check assumptions
plot(RSR_model_3)
```

### Check for Nonsignificant Interactions
```{r}
anova(RSR_model_3)
```
**Final Model: 2-way with 2 interaction effect**

## Final Model: Root Shoot Ratio = Wheat + Soil + Biochar + Wheat:Soil + Soil:Biochar
```{r}
interaction.plot(Soil, Biochar, root_shoot_ratio)
interaction.plot(Soil, Wheat, root_shoot_ratio)
```

### Multiple Comparison
```{r}
# compare the effect of Wheat at each level of Soil 
emmeans(RSR_model_3, list(pairwise ~ Wheat|Soil)) 
```
+ The effect of Wheat is significant when Soil type is "soil-sand" (p-value = 0.0011 < 0.05) in terms of root shoot ratio.

```{r}
# compare the effect of Soil at each level of Wheat
emmeans(RSR_model_3, list(pairwise ~ Soil|Wheat))
```
+ The effect of Soil is significant at each level of Wheat type (both p-value < 0.05) in terms of root shoot ratio.

```{r}
# compare the effect of Wheat at each level of Soil 
emmeans(RSR_model_3, list(pairwise ~ Soil|Biochar)) 
```
+ The effect of Soil is significant when at each level of Biochar treatment (both p-value < 0.05) in terms of root shoot ratio.

```{r}
# compare the effect of Soil at each level of Wheat
emmeans(RSR_model_3, list(pairwise ~ Biochar|Soil))
```
+ The effect of Biochar is significant when Soil type is "soil" (p-value = 0.0002 < 0.05) in terms of root shoot ratio.















